### 2025-01-24 isabelle/CVM_Subsample/document/root.tex
slightly edit abstract

\documentclass[11pt,a4paper]{article}
\usepackage{isabelle,isabellesym}
\usepackage{amsfonts,amsmath, amssymb}
\usepackage[top=1.4in, bottom=1.4in, left=1.4in, right=1.4in]{geometry}

\usepackage[only,bigsqcap]{stmaryrd}

\usepackage{algorithm}
\usepackage[noend]{algpseudocode}
\renewcommand{\algorithmicrequire}{\textbf{Input:}}
\renewcommand{\algorithmicensure}{\textbf{Output:}}
\let\oldsection\section
\renewcommand\section{\clearpage\oldsection}

\newcommand{\getsr}{\xleftarrow{\$}}

\usepackage{mathtools}
\DeclarePairedDelimiter{\ceil}{\lceil}{\rceil}
\DeclareMathOperator{\prob}{\mathcal P}
\DeclareMathOperator{\expect}{\mathbb E}
\DeclareMathOperator{\Ber}{\mathrm{Ber}}
\DeclareMathOperator{\indicator}{\mathrm{I}}

% this should be the last package used
\usepackage{pdfsetup}

% urls in roman style, theory text in math-similar italics
\urlstyle{rm}
\isabellestyle{it}

\begin{document}

\title{Verification of the CVM algorithm with a new Recursive Analysis Technique}
\author{Emin Karayel, Derek Khu, Kuldeep S. Meel, Yong Kiam Tan,\\and Joe Watt}
\maketitle

\begin{abstract}
In 2022, Chakraborty et al.~\cite{chakraborty2022} published a streaming
algorithm (henceforth, the CVM algorithm) for the distinct elements problem
that deviated considerably from the state-of-the art, due to its simplicity
and avoidance of standard derandomization techniques, while still maintaining
a close-to-optimal logarithmic space complexity.

In this entry, we verify the CVM algorithm's correctness using a new technique
which simplifies the formal analysis considerably compared to the original
proof by Chakraborty et al.. The main idea is based on the well-known notion of
probabilistic recurrence relations. However, instead of recurrence relations
with respect to the state, we investigate general functions of the state and
derive conditions on the function which preserve the relevant recurrence
relation. This generalization allows us to derive concentration bounds which is
usually not possible with recurrence relations.

This new technique opens up the possible algorithm design space, and we
introduce a new variant of the CVM algorithm that uses less space, fewer steps,
and also has an additional property in addition to concentration: unbiasedness.
The latter property means that the new variant returns the correct distinct
element count in expectation, which is a property that neither the original CVM
algorithm nor classic algorithms for the distinct elements problem posses.
% TODO(YK): in what sense does it use less space and fewer steps?
\end{abstract}

\tableofcontents

\parindent 0pt\parskip 0.5ex

\input{session}

\bibliographystyle{abbrv}
\bibliography{root}

\end{document}