### 2025-02-19 paper/sec/abstract.tex
Paper: Add more details on how the invariant is used in Sec 3.   Minor edits: Avoid ``transformation-based`` in abstract (which requires additional context)   Section 4 (clarify range of S)

\begin{abstract}
Estimating the number of distinct elements in a data stream is a classical problem with numerous applications in computer science.
We formalize a recent, remarkably simple, randomized algorithm for this problem due to Chakraborty, Vinodchandran, and Meel (called CVM algorithm).
Their algorithm deviated considerably from the state-of-the art, due to its avoidance of intricate derandomization techniques, while still maintaining a close to optimal logarithmic space complexity.

Central to our formalization is a new proof technique based on \emph{functional probabilistic invariants}, which allows us to derive concentration bounds using the Cram\'{e}r--Chernoff method.
This simplifies the formal analysis considerably compared to the original proof by Chakraborty et al.
Moreover, our technique opens up the possible algorithm design space; we demonstrate this by introducing and verifying with a new variant of the CVM algorithm that is total and unbiased---neither property is possessed by the original algorithm.
In this paper, we introduce the proof technique, describe its use to mechanize both versions of the CVM algorithm in Isabelle/HOL, and present a supporting formalized library on negatively associated random variables used to verify the latter variant.
\end{abstract}