### 2025-02-27 isabelle/CVM_Transforms/document/root.tex
wip docs

\documentclass[11pt, a4paper]{report}
\usepackage[T1]{fontenc}
\usepackage{isabelle, isabellesym, amsmath}

% this should be the last package used
\usepackage{pdfsetup}

% urls in roman style, theory text in math-similar italics
\urlstyle{rm}
\isabellestyle{it}

\begin{document}

\title{Submission Example}
\author{placeholder}
\maketitle

\begin{abstract}
  This is an example submission to the Archive of Formal Proofs. It
  shows submission requirements and explains the structure of a simple
  typical submission.
\end{abstract}

\tableofcontents

% include generated text of all theories
\input{session}

% \bibliographystyle{abbrv}
% \bibliography{root}

\end{document}