### 2025-02-28 isabelle/CVM_Transforms/document/root.tex
condense theories, add docs

\documentclass[11pt, a4paper]{report}

\usepackage[T1]{fontenc}
\usepackage{isabelle, isabellesym}
\usepackage{amsfonts, amsmath, amssymb}

% this should be the last package used
\usepackage{pdfsetup}

% urls in roman style, theory text in math-similar italics
\urlstyle{rm}
\isabellestyle{it}

\begin{document}

\title{Verification of the CVM algorithm}

\author{
  Seng Joe Watt,
  Derek Khu,
  Emin Karayel,
  Kuldeep S. Meel,
  \\
  and Yong Kiam Tan}

\maketitle

\begin{abstract}
  This is an example submission to the Archive of Formal Proofs. It
  shows submission requirements and explains the structure of a simple
  typical submission.
\end{abstract}

\tableofcontents

\parindent 0pt
\parskip 0.5ex

\chapter{CVM algorithm formalisation}

\input{CVM_Algo}
\input{CVM_Algo_No_Fail}
\input{CVM_Algo_Lazy_Eager}
\input{CVM_Algo_Nondet_Binomial}

\input{CVM_Error_Bounds}
\input{CVM_Correctness}
\input{CVM_Correctness_Instance}

\chapter{Utility definitions and results}

\input{Utils_Basic}
\input{Utils_Reader_Monad}
\input{Utils_Approx_Algo}
\input{Utils_Concentration_Ineqs}

\input{Utils_PMF_Basic}
\input{Utils_PMF_Lazify}

\input{Utils_SPMF_Basic}
\input{Utils_SPMF_Rel_Hoare}

\chapter{References}

\bibliographystyle{abbrv}
\bibliography{root}

\end{document}