### 2025-03-04 isabelle/CVM_Transforms/document/root.tex
slight fixes

\documentclass[11pt, a4paper]{report}

\usepackage[T1]{fontenc}
\usepackage{isabelle, isabellesym}
\usepackage{amsfonts, amsmath, amssymb}
\usepackage[htt]{hyphenat}

% this should be the last package used
\usepackage{pdfsetup}

% urls in roman style, theory text in math-similar italics
\urlstyle{rm}
\isabellestyle{it}

\begin{document}

\title{Verification of the CVM algorithm}

\author{
  Seng Joe Watt,
  Derek Khu,
  Emin Karayel,
  Kuldeep S. Meel,
  \\
  and Yong Kiam Tan
}

\maketitle

\begin{abstract}
  Published in 2022, the CVM algorithm \cite{cvm_2022} is a remarkably simple
  one for the distinct elements problem, avoiding intricate derandomization
  techniques, while maintaining a close to optimal logarithmic space complexity.
  Later, it was discovered that the proof was incorrect, with the authors
  presenting a new proof in \cite{cvm_2023}.

  This project formalises the new, corrected proof of the CVM algorithm, while
  strengthening the correctness result of Theorem 2 in \cite{cvm_2023} to provide
  the same correctness guarantee using a slightly smaller threshold
  (and hence space usage) of
  $\lceil \frac{12}{\varepsilon^2} \text{log}_2 \frac{3m}{\delta} \rceil$,
  where $m$ is the length of the input stream.

  This proof performs a sequence of transformations from the CVM algorithm,
  akin to cryptographic game-hopping.  The reason for these transformations is
  that standard techniques for analysing randomised algorithms such as the
  Chernoff bounds usually require independence of state random variables, which
  is unfortunately absent from the CVM algorithm.  The transformations address
  this by successively simplifying the algorithm into one where the state
  variables are independent, enabling the use of the Chernoff bounds.

  As part of the project, we also formalise various useful results, in particular:
  \begin{enumerate}
    \item Unary and relational Hoare rules for reasoning about monadic folds
    (ie. loops) over the \texttt{reader}, \texttt{pmf} and \texttt{spmf} monads,
    which we use to justify the various program transformations.

    \item Strong forms of the multiplicative Chernoff bounds for the binomial
    distribution, derived from the Bennett-Bernstein inequalities.
  \end{enumerate}
\end{abstract}

\tableofcontents

\parindent 0pt
\parskip 0.5ex

\chapter{CVM algorithm formalisation}

\input{CVM_Algo}
\input{CVM_Algo_No_Fail}
\input{CVM_Algo_Lazy_Eager}
\input{CVM_Algo_Nondet_Binomial}

\input{CVM_Error_Bounds}
\input{CVM_Correctness}
\input{CVM_Correctness_Instance}

\chapter{Utility definitions and results}

\input{Utils_Basic}
\input{Utils_Reader_Monad}
\input{Utils_Approx_Algo}
\input{Utils_Concentration_Ineqs}

\input{Utils_PMF_Basic}
\input{Utils_PMF_Lazify}

\input{Utils_SPMF_Basic}
\input{Utils_SPMF_Rel_Hoare}

\bibliographystyle{abbrv}
\bibliography{root}

\end{document}