### 2025-06-20 paper/sec/conclusion.tex
Clarify that the novel part of the new technique is that we are establishing exponential tail bounds.

\section{Conclusion}\label{sec:conclusion}
We presented the first formalization of the CVM algorithm using Isabelle/HOL.
Central to our formalization is a novel invariant-based proof technique to establish exponentially decreasing tail-bounds for randomized algorithms, which is inspired by our alternative analysis of the CVM algorithm via the Cram\'{e}r--Chernoff method; comparing our approach against the original proof by Chakraborty et al.~\cite{chakraborty2023} shows that our technique yields a considerably shorter formalization (with \locnew~vs.~\locold~lines).
Interestingly, our technique also readily generalized to a new CVM variant with stronger properties (totality and unbiasedness)---we formalized this latter version using the same invariant, together with a new library of results for negative association.

In future work, it would be interesting to formalize other variations of subsampling for CVM.
More generally, one could further explore whether the technique we introduced here could be applied towards proofs of other randomized (streaming) algorithms.