### 2025-09-19 presentation/algorithm_animation.tex
Presentation Draft

%\newcommand{streamcolor}{#5D8AA8}
\definecolor{streamcolor1}{rgb}{1.0, 0.94, 0.0}
%\definecolor{streamcolor1}{rgb}{0.36, 0.54, 0.66}
\definecolor{streamcolor2}{rgb}{0.9, 0.17, 0.31}
\definecolor{streamcolor3}{rgb}{0.0, 0.5, 0.0}
\definecolor{streamcolor4}{rgb}{0.0, 0.5, 1.0}
\definecolor{streamcolor5}{rgb}{0.74, 0.2, 0.64}
\definecolor{tailcoinbordercolor}{rgb}{0.13, 0.67, 0.8}
\definecolor{tailcoinfillcolor}{rgb}{0.54, 0.81, 0.94}
\definecolor{headcoinbordercolor}{rgb}{1.0, 0.44, 0.37}
\definecolor{headcoinfillcolor}{rgb}{0.98, 0.91, 0.71}

\newcommand{\illustratestate}[7]{
  \begin{tikzpicture}[
    streamcell/.style={circle, draw=black, fill=black!15, thick, text centered,inner sep=3pt,minimum width=0.6cm},
    arrow/.style={thick,-{Latex[length=.32cm, width=.32cm,fill=white]}},
    coin/.style={circle, draw=orange, fill=yellow, thick},
  ]
  \node[streamcell,fill=streamcolor3] at (0cm,0 cm) {};
  \node[streamcell,fill=streamcolor1] at (1cm,0 cm) {};
  \node[streamcell,fill=streamcolor3] at (2cm,0 cm) {};
  \node[streamcell,fill=streamcolor2] at (3cm,0 cm) {};
  \node[streamcell,fill=streamcolor4] at (4cm,0 cm) {};
  \node[streamcell,fill=streamcolor5] at (5cm,0 cm) {};
  \node[streamcell,fill=streamcolor1] at (6cm,0 cm) {};
  \node[streamcell,fill=streamcolor2] at (7cm,0 cm) {};
  \node[streamcell,fill=streamcolor3] at (8cm,0 cm) {};
  \node at (9cm,0cm) {$\cdots$};
  \draw[arrow] (1cm * #1, -1.5cm) to (1cm * #1, -0.5cm);
  \node (robot) at (0,-3.5cm)
    {\includegraphics[width=.25\textwidth]{contrib/vecteezy_cute-robot-holding-clipboard-cartoon-icon-illustration_53779545_655/vecteezy_cute-robot-holding-clipboard-cartoon-icon-illustration_53779545}};

  \node[anchor=west] at (1.4cm, -3.1cm) {$\chi = \{$};
  \ifthenelse{\equal{#3}{empty}}{}{\node[streamcell,fill=#3] at (3cm, -3.1cm) {};}
  \ifthenelse{\equal{#4}{empty}}{}{\node[streamcell,fill=#4] at (3.8cm, -3.1cm) {};}
  \ifthenelse{\equal{#5}{empty}}{}{\node[streamcell,fill=#5] at (4.6cm, -3.1cm) {};}
  \ifthenelse{\equal{#6}{empty}}{}{\node[streamcell,fill=#6] at (5.4cm, -3.1cm) {};}
  \node at (6cm, -3.1cm) {$\}$};

  \ifthenelse{\equal{#7}{head}}{
    \node[coin,fill=headcoinfillcolor,draw=headcoinbordercolor] at (-1cm,-3.5cm) {H}; 
  }{}
  \ifthenelse{\equal{#7}{tail}}{
    \node[coin,fill=tailcoinfillcolor,draw=tailcoinbordercolor] at (-1cm,-3.5cm) {T}; 
  }{}

  \node[anchor=west] at (1.4cm, -3.8cm) {$p = #2, n = 4$};

  \end{tikzpicture}
}

\begin{frame}
\frametitle{Illustration of the Algorithm}

\only<1>{\illustratestate{0}{1}{empty}{empty}{empty}{empty}{none}}
\only<2>{\illustratestate{0}{1}{empty}{empty}{empty}{empty}{head}}
\only<3>{\illustratestate{0}{1}{streamcolor3}{empty}{empty}{empty}{head}}
\only<4>{\illustratestate{1}{1}{streamcolor3}{empty}{empty}{empty}{head}}
\only<5>{\illustratestate{2}{1}{streamcolor3}{streamcolor1}{empty}{empty}{head}}
\only<6>{\illustratestate{3}{1}{streamcolor3}{streamcolor1}{empty}{empty}{head}}
\only<7>{\illustratestate{4}{1}{streamcolor3}{streamcolor1}{streamcolor2}{empty}{head}}
\only<8>{\illustratestate{4}{1}{streamcolor3}{streamcolor1}{streamcolor2}{streamcolor4}{head}}
\only<9>{\illustratestate{4}{1}{empty}{streamcolor1}{streamcolor2}{empty}{none}}
\only<10>{\illustratestate{4}{\frac{1}{2}}{empty}{streamcolor1}{streamcolor2}{empty}{none}}
\only<11>{\illustratestate{5}{\frac{1}{2}}{empty}{streamcolor1}{streamcolor2}{empty}{none}}
\only<12>{\illustratestate{5}{\frac{1}{2}}{empty}{streamcolor1}{streamcolor2}{empty}{head}}
\only<13>{\illustratestate{5}{\frac{1}{2}}{streamcolor5}{streamcolor1}{streamcolor2}{empty}{head}}
\only<14>{\illustratestate{6}{\frac{1}{2}}{streamcolor5}{streamcolor1}{streamcolor2}{empty}{none}}
\only<15>{\illustratestate{6}{\frac{1}{2}}{streamcolor5}{streamcolor1}{streamcolor2}{empty}{tail}}
\only<16>{\illustratestate{6}{\frac{1}{2}}{streamcolor5}{empty}{streamcolor2}{empty}{taiil}}
\only<17>{\illustratestate{7}{\frac{1}{2}}{streamcolor5}{empty}{streamcolor2}{empty}{taiil}}
\end{frame}