### 2025-09-27 presentation/main.tex
Minor edits and split into sections.

\documentclass[xcolor=table]{beamer}
\usetheme{Rochester}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{tikz}
\usepackage{pgfplots}
\pgfplotsset{compat=1.18}
\usetikzlibrary{arrows.meta}
\usetikzlibrary{calc}
\usetikzlibrary{decorations.pathmorphing}
\newcommand*\circled[1]{\tikz[baseline=(char.base)]{
            \node[shape=circle,draw, minimum size=3.5mm,inner sep=0.5pt] (char) {#1};}}

\usepackage{ifthen}
\usepackage[noend]{algpseudocode}
\usepackage{soul}
\usepackage{isabelle}
\usepackage{isabellesym}
\isabellestyle{it}

\DeclareMathOperator{\mbind}{\text{\isa{\isasymbind}}}
\DeclareMathOperator{\expect}{\mathrm{E}}

\renewcommand{\algorithmicrequire}{\textbf{Input:}}
\renewcommand{\algorithmicensure}{\textbf{Output:}}
\newcommand{\ift}[3]{\mathbf{if} \; #1 \; \mathbf{then} \; #2 \; \mathbf{else} \; #3}
\newcommand{\integral}[3]{\int_{#1} \! #2 \, \mathrm{d} #3}

\newcommand{\getsr}{\xleftarrow{\$}}
\newcommand{\robot}[1]{\includegraphics[width=#1]{contrib/vecteezy_cute-robot-holding-clipboard-cartoon-icon-illustration_53779545_655/vecteezy_cute-robot-holding-clipboard-cartoon-icon-illustration_53779545}}

\title{Verification of the CVM algorithm with a Functional Probabilistic Invariant}
\author{E. Karayel\inst{1}, J. Watt\inst{2}, D. Khu\inst{2}, K. Meel\inst{4,5}, Y. K. Tan\inst{2,3}}
\institute[VFU]{
  \inst{1}%
    Technical University of Munich, Germany
  \and
  \inst{2}%
    Institute for Infocomm Research (I$^2$R), A*STAR, Singapore
  \and
  \inst{3}%
    Nanyang Technological University, Singapore
  \and
  \inst{4}%
    Georgia Institute of Technology, Atlanta, GA, USA
  \and
  \inst{5}%
    University of Toronto, Canada
}

\usepackage{mathtools}
\DeclareMathOperator{\prob}{\mathcal P}
\DeclareMathOperator{\bigo}{\mathcal O}
\DeclarePairedDelimiter{\ceil}{\lceil}{\rceil}
\DeclareMathOperator{\Ber}{\mathrm{Ber}}
\DeclareMathOperator{\indicat}{\mathrm{I}}

\begin{document}
\AtBeginSection[]
{
    \begin{frame}
        \frametitle{Table of Contents}
        \tableofcontents[currentsection]
    \end{frame}
}

\frame{\titlepage}
\begin{frame}
\frametitle{Table of Contents}
\tableofcontents
\end{frame}

\input{sections/intro.tex}

\input{sections/prob_fun_invariants.tex}

\input{sections/original_proof.tex}

\input{sections/conclusion.tex}

\end{document}