### 2025-09-27 presentation/sections/conclusion.tex

\begin{frame}
\frametitle{Conclusion/Summary}
\begin{itemize}
\item Verification of the CVM Algorithm
\pause
\item New Technique for the Verfication of Randomized Algorithms
\pause
\item New unbiased/total variant of the CVM Algorithm
\pause
\item Development of a Library for Negative Association
\pause
\begin{itemize}
\item A generalization of the concept of independence
\pause
\item Example: Indicator random variables of an idealized Bloom-Filter
\end{itemize}
\pause
\item Thank You!
\pause
\item Artwork: \raisebox{-0.3cm}{\robot{1cm}} By Vecteezy.com.
\end{itemize}
\end{frame}

\section*{References}
\begin{frame}[allowframebreaks]
        \frametitle{References}
        \bibliographystyle{plain}
        \bibliography{main.bib}
\end{frame}