### 2025-09-27 presentation/sections/original_proof.tex

\section{Sketch of the Original Proof}
\begin{frame}
\frametitle{Original Proof}
\begin{tikzpicture}[node distance=2.7cm, auto,scale=0.8,every node/.style={scale=0.8}]

    \node (A) [draw, rectangle, minimum width=2.1cm, minimum height=1.2cm, align=center] {CVM};
    \node (B) [draw, rectangle, minimum width=2.1cm, minimum height=1.2cm, align=center, right of=A] {CVM w/o\\ Step 3};
    \node (C) [draw, rectangle, minimum width=2.1cm, minimum height=1.2cm, align=center, right of=B] {Eager Alg.};
    \node (D) [draw, rectangle, minimum width=2.1cm, minimum height=1.2cm, align=center, right of=C] {Random \\ Process};
    \node (E) [draw, rectangle, minimum width=2.1cm, minimum height=1.2cm, align=center, right of=D] {Chernoff \\ Bounds};

    \node (C') [draw, rectangle, minimum width=2.1cm, minimum height=1.2cm, align=center, below of=B, node distance = 2.0cm] {Functional\\Invariant};

    \node (pac) [draw, rectangle, minimum width=2.1cm, minimum height=1.2cm, align=center, below of=E, node distance = 2.0cm] {CVM PAC\\ Guarantee};

    \draw[thick] ($(A.north west) + (-0.3,0.2)$) rectangle ($(E.south east) + (0.3,-0.6)$);
    \node[anchor=south] at ($(D.south) + (0.0,-0.6)$) {Transformation-based Proof};

    \draw[dashed,thick] ($(A.north west) + (-0.5,0.4)$) rectangle ($(C'.south east) + (0.2,-0.2)$);
    \node[anchor=south,align=center] at ($(C'.west) + (-1.5,-0.45)$) {Probabilistic\\Invariant Proof};

    \draw[->] (A) -- (B);
    \draw[->] (B) -- (C);
    \draw[->] (B) -- (C');
    \draw[->] (C) -- (D);
    \draw[->] (D) -- (E);
    \draw[->] (E) -- (pac);
    \draw[->] (C') -- (pac);

\end{tikzpicture}

\end{frame}

\newcommand{\overlaytransform}[1]{
\begin{tikzpicture}[node distance=2.7cm, auto,scale=0.8,every node/.style={scale=0.8}]

\ifthenelse{#1=1}{\colorlet{stagecolor}{gray}}{\colorlet{stagecolor}{white}}
\node (A) [draw, rectangle, minimum width=2.1cm, minimum height=1.2cm, align=center, fill=stagecolor] {CVM};
\ifthenelse{#1=2}{\colorlet{stagecolor}{gray}}{\colorlet{stagecolor}{white}}
\node (B) [draw, rectangle, minimum width=2.1cm, minimum height=1.2cm, align=center, fill=stagecolor, right of=A] {CVM w/o\\ Step 3};
\ifthenelse{#1=3}{\colorlet{stagecolor}{gray}}{\colorlet{stagecolor}{white}}
\node (C) [draw, rectangle, minimum width=2.1cm, minimum height=1.2cm, align=center, fill=stagecolor, right of=B] {Eager Alg.};
\ifthenelse{#1=4}{\colorlet{stagecolor}{gray}}{\colorlet{stagecolor}{white}}
\node (D) [draw, rectangle, minimum width=2.1cm, minimum height=1.2cm, align=center, fill=stagecolor, right of=C] {Random \\ Process};
\draw[->] (A) -- (B);
\draw[->] (B) -- (C);
\draw[->] (C) -- (D);
\draw[thick] ($(A.north west) + (-0.3,0.2)$) rectangle ($(D.south east) + (0.3,-0.3)$);
\end{tikzpicture}
}

\begin{frame}
\frametitle{CVM}
\overlaytransform{1}
\begin{algorithmic}
  \State $\chi \gets \{\}, p \gets 1, n = \ceil*{\frac{12}{\varepsilon^2} \ln{(\frac{6 l}{\delta})} }$
  \For{$i \gets 1$ to $l$}
    \State $b \getsr \Ber(p)$
    \If{$b$}
      \State $\chi \gets \chi \cup \{a_i\}$
    \Else
      \State $\chi \gets \chi - \{a_i\}$
    \EndIf
    \If{$|\chi| = n$}
      \State $\chi \getsr \mathrm{subsample}(\chi)$
      \State $p \gets \frac{p}{2}$
    \EndIf
    \If{$|\chi| = n$}
      \Return $\bot$
    \EndIf
  \EndFor
  \State \Return $\frac{|\chi|}{p}$
\end{algorithmic}
\end{frame}

\begin{frame}
\frametitle{CVM Algorithm II}
\overlaytransform{2}
\begin{algorithmic}
  \State $\chi \gets \{\}, k \gets 0, n = \ceil*{\frac{12}{\varepsilon^2} \ln{(\frac{6l}{\delta})} }$
  \For{$i \gets 1$ to $l$}
    \State $b \getsr \Ber(2^{-k})$ 
    \If{$b$} 
      \State $\chi \gets \chi \cup \{a_i\}$
    \Else 
      \State $\chi \gets \chi - \{a_i\}$
    \EndIf
    \If{$|\chi| = n$} 
      \State $\chi \getsr \mathrm{subsample}(\chi)$
      \State $k \gets k + 1$
    \EndIf
  \EndFor
  \State \Return $2^k |\chi|$ 
\end{algorithmic}
\end{frame}

\begin{frame}
\frametitle{Eager Algorithm}
\overlaytransform{3}
\begin{algorithmic}
  \State $\chi \gets \{\}, k \gets 0, n = \ceil*{\frac{12}{\varepsilon^2} \ln{(\frac{6l}{\delta})} }$
  \only<2->{\color{red}}
  \State $b[i,j] \getsr \Ber(1/2)$ for $i,j \in \{1,\cdots,l\}$ % \Comment perform $l^2$ unbiased independent coin flips
  \only<2->{\color{black}}
  \For{$i \gets 1$ to $l$}
    \If{$b[i,1]=b[i,2]=\cdots=b[i,k]=1$}
      \State $\chi \gets \chi \cup \{a_i\}$
    \Else
      \State $\chi \gets \chi - \{a_i\}$
    \EndIf
    \If{$|\chi| = n$} % \Comment if buffer $\chi$ is full
      \State $\chi \gets \{a \in \chi \;|\; b[\textrm{last{\textunderscore}index}(a),k+1] = 1\}$ % \Comment keep elems.~whose $k+1$-th flip is $1$
      \State $k \gets k+1$
    \EndIf
  \EndFor
  \State \Return $2^k |\chi|$ 
  \end{algorithmic}
\end{frame}

\begin{frame}
\frametitle{Random Process}
\overlaytransform{4}

\begin{block}{Observation}
The eager algorithm preserves the invariant:
\[
  s \in \chi \leftrightarrow \text{For all $j < k$:} b[ \textrm{last{\textunderscore}index}(s), j] = 1
\]
\end{block}
\pause
Non-deterministic algorithm:
\begin{algorithmic}
  \State $b[i,j] \getsr \Ber(1/2)$ for $i,j \in \{1,\cdots,l\}$ % \Comment perform $l^2$ unbiased independent coin flips
  \State $k \gets \{0,\dots,k_\mathrm{max}\}$ \Comment Non-deterministc step.
  \State $\chi \gets \{ s \in A \mid b[ \textrm{last{\textunderscore}index}(s), j] = 1 \textrm { for all } j < k \}$
  \State \Return $2^k |\chi|$ 
\end{algorithmic}
\end{frame}